---
title: "Live Session Unit 04"
author: "Claudia Woodruff"
date: "September 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: FiveThirtyEight Data found at https://github.com/rudeboybert/fivethirtyeight

#### 1a) Install the fivethirtyeight package

```{r}
#install.packages("fivethirtyeight")
library("fivethirtyeight")
```

#### 1b) . In the listing of Data sets in package ‘fivethirtyeight,’ assign the eighteenth data set toan object ‘df.’

```{r getEighteenth }
d <- data(package="fivethirtyeight")
nm <- d$results[18,"Item"]
df <- college_recent_grads
```

#### 1C) URL to related news
Use a more detailed list of the data sets to write out the URL in a comment to the related news story.
The URL below was obtained by:
**vignette("fivethirtyeight", package = "fivethirtyeight")**

http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/

#### 1d) Using R command(s), give the dimensions and column names of this data frame.

```{r dimcol }
dim(df)
names(df)

```


## Question 2: Use your newly assigned data frame for Question 2.
#### 2a) Column names and counts
Write an R command that gives you the column names of the data frame. Right after that, write one that counts the number of columns but not rows. 

```{r countCol }
names(df)
ncol(df)
```

#### 2b) Count each major categories.
Generate a count of each unique major_category in the data frame. Assign it to major_count.

```{r maj_cat_count }
library(plyr)
major_count <- count(college_recent_grads, "major_category")
major_count
```

#### 2c) In a single pass of barplot(), make a barplot of major_count.
To make things easier to read, enter par(las=2) before your plot to make the text perpendicular to the axis. Make sure to label the title with something informative, label the x and y axis, and make it any color other than grey. Assign the major_category labels to their respective bar. Flip the barplot horizontally so that bars extend to the right, not upward.

```{r plotCount }
# Fitting Labels 
par(las=2)          # make label text perpendicular to axis
par(mar=c(5,15,4,2)) # increase y-axis margin.


barplot(major_count$freq, 
        main="Frequency of Majors by Categories",
        horiz=TRUE, 
        names.arg=major_count$major_category,
        xlab="Frequency",
        col="green")

```

#### 2d) Write the fivethirtyeight data to a csv file

```{r write_to_file}
write.csv(df, file = "CollegeRecentGrads.csv", row.names=FALSE)
```

## Question 3 GitHub Repository with Unit04 Assignments
##### (RMarkdown script, HTML file, and data from ‘fivethirtyeight.’ )

https://github.com/cruffwood/Homework.git
