---
title: "Live Session Unit 05 - Tidy Data"
author: "Claudia Woodruff"
date: "October 4, 2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1: Data Munging utilizing **yob2016.txt**

#### 1a) Import the .txt file into R so you can process it.

```{r Import_2016 } 

df <- read.table("yob2016.txt", sep = ";", header = FALSE)
colnames(df) <- c("FName", "Sex", "Y2016Amt")
colnames(df)
         
```

#### 1b) Display the summary and structure of df.

```{r Disp_df}

summary(df)
str(df)

```

#### 1c/d) Remove the observation where the Name has "yyy" at the end.
####       and write an R command to figure out which name it is and display it.

```{r FindName}
# Find the name entered twice with the three y's and display it.
yyy <- df$FName[grep("yyy", df$FName)]
yyy

y2016 <- df[-which(as.character(df$FName) == df$FName[grep("yyy", df$FName)]),]
str(y2016)

```


## Question 2: Data Merging utilizing yob2015.txt
##### This file is similar to yob2016, but contains names, gender, and total children given that name for the year **2015**.

#### 2a)Import the .txt file into R. 


```{r Import_2015}

y2015 <- read.csv("yob2015.txt", header = FALSE)
colnames(y2015) <- c("FName", "Sex", "Y2015Amt")
str(y2015)
summary(y2015)

```

#### 2b) Display the last ten rows in the dataframe. Describe something you find interesting about these 10 rows.

#####     The names appear to be in ascending order. Both FName and Sex are factors.

```{r }

tail(y2016,10)

```

#### 2c) Merge y2016 and y2015 by your Name column; assign it to final. 
* The client only cares about names that have data for both 2016 and 2015.
* There should be no NA values in either of your amount of children rows after merging.

```{r }
                       
final <- merge(x = y2016, y = y2015)       # Will match on "FName" and "Sex"

summary(final)

```

## Question 3: Data Summary

#### 3a) Add the amount of children in 2015 and 2016 in a new column called "Total".
In those two years combined, how many people were given popular names? 

```{r }
# Add amounts 2015 and 2015 into 'final'
final$Total <- rowSums(final[,c("Y2015Amt", "Y2016Amt")])

str(final)  # 26550 observations

```

#### 3b) Sort the data by Total. What are the top 10 most popular names?

```{r }

final <- final[order(final$Total, decreasing = TRUE), ]
head(final, 10)

```

####  3c) The client is expecting a girl! Omit boys and give the top 10 most popular girl's names. Assign this to object **girl**.

```{r GirlsOnly}

girl <- final[final$Sex %in% c("F"), ] 

head(girl, 10)

```

####  3d) Write these top 10 girl Names and their Totals to a CSV file called **itsagirl.csv**. 
* Do not include row labels. 
* Leave out columns other than Name and Total.

```{r }

itsagirl <-girl[1:10, c("FName", "Total")]
write.csv(itsagirl, file = "itsagirl.csv", row.names = FALSE)

```

## Question 4: Upload to GitHub
### https://github.com/cruffwood/Homework

